---
title: "Stat 435 Homework 2"
author: "Brian Bruxvoort"
date: "2023-09-24"
output:
  bookdown::pdf_document2:
    latex_engine: lualatex
  tufte::tufte_handout:
    latex_engine: xelatex
---

# Page 52, Number 14, part a
Pi is between 0 and 1, so the range for log(pi/(1-pi)) is between 0 and 1 after the end points back using the exp(.)/[1 + exp(.)] transformation. As you can see the new numerical range is the same as before because log and exp cancel out. The reason this range is ideal is because it gives us a probability of an event happening between 0 and 1, which is exactly what we want.

# Page 52, Number 14, part b
Verify the estimated variance of log (ˆπ/(1 − ˆπ)) is [nˆπ(1 − ˆπ)]−1 by using the
delta method (see Appendix B.4.2).
```{r}
pi.hat <- c(0.6, 0.65, 0.7, 0.75, 0.8)
g <- log(pi.hat/(1-pi.hat))
g
var(g)

n <- length(pi.hat)
h <- (n*pi.hat*(1 - pi.hat))-1
sum(h)
```

# Page 52, Number 14, part c
```{r}
install.packages("binom")
library(binom)
n <- 10
w <- 4
conf_interval <- binom.confint(w, n, method = "logit", conf.level = 0.95)
conf_interval
```

# Page 54, Number 17, part a
```{r}
c.table <- array(data = c (22, 10, 4, 6) , dim = c(2, 2),
dimnames = list("Strategy" = c("No time-out", "Time-out"),
"Field Goal" = c( "Success", "Failure")))
c.table
```

```{r}
#Wald
pi.hat.table <- c.table/rowSums(c.table)
alpha <- 0.05
pi.hat1 <- pi.hat.table[1,1]
pi.hat2 <- pi.hat.table[2,1]
var.wald <- pi.hat1 * (1 - pi.hat1)/sum(c.table[1 ,]) +
pi.hat2 * (1 - pi.hat2) / sum(c.table [2 ,])
pi.hat1 - pi.hat2 + qnorm(p = c(alpha / 2, 1- alpha / 2))*
sqrt(var.wald)
```

In the Wald confidence interval, the interval contains 0, which means there is not sufficient evidence to indicate a difference in using a time out to "ice" a kicker. That means it doesn't matter if you use a time out.

```{r}
#Agresti-Caffo
pi.tilde1 <- (c.table[1 ,1] + 1)/(sum(c.table[1 ,]) + 2)
pi.tilde2 <- (c.table[2 ,1] + 1)/(sum(c.table[2 ,]) + 2)
var.AC <- pi.tilde1 * (1 - pi.tilde1)/(sum(c.table [1,]) + 2) + pi.tilde2 * (1 - pi.tilde2)/(sum(c.table [2 ,]) + 2)
pi.tilde1 - pi.tilde2 + qnorm(p = c(alpha / 2, 1-alpha /2))*sqrt(var.AC)
```

For the Agresti-Caffo confidence interval, the interval also contains 0, which means there is not sufficient evidence to indicate a difference in using a time out to "ice" a kicker either. Both models agree, so that means it really doesn't matter if you use a time-out to "ice" a kicker, statistically, they have the same probability of making it.

# Page 54, Number 17, part b
```{r}
library(package = PropCIs)
prop.test(x = c.table, conf.level = 0.95, correct = FALSE)
```

```{r, warning=FALSE}
chisq.test(x = c.table, correct = FALSE)
```


# Page 54, Number 17, part c
```{r}
pi.hat.table <- c.table/rowSums(c.table)
pi.hat.table
```

```{r}
pi.hat1 <- pi.hat.table[1,1]
pi.hat2 <- pi.hat.table[2,1]

round(pi.hat1/pi.hat2, 4)

round(1/(pi.hat1/pi.hat2), 4) #inverted

alpha<-0.05
n1 <- sum(c.table[1,])
n2 <- sum(c.table[2,])
var.log.rr <- (1-pi.hat1) / (n1*pi.hat1) + (1-pi.hat2) / (n2*pi.hat2)
var.log.rr
```

```{r}
ci <- exp(log(pi.hat1/pi.hat2) + qnorm(p = c(alpha/2, 1-alpha/2)) * sqrt(var.log.rr))
round(ci, 4)

rev(round(1/ci, 4)) #inverted
```

The relative risk inverted is 0.7386, which means it is 27% less likely that a kicker will make a field goal when he is iced. With 95% confidence, we can say a kicker will be less likely to make a field goal between 51% and -12% of the time when he is iced. It is a pretty wide interval so that is something to keep an eye on. 

# Page 54, Number 17, part d
```{r}
OR.hat <- c.table[1,1] * c.table[2,2] / (c.table[2,1] * c.table[1,2])
round(OR.hat, 4)

round(1/OR.hat, 4)
```

```{r}
alpha <- 0.05
var.log.or <- 1/c.table[1,1] + 1/c.table[1,2] + 1/c.table[2,1] + 1 /c.table[2,2]
OR.CI <- exp(log(OR.hat) + qnorm(p = c(alpha / 2, 1-alpha / 2)) * sqrt(var.log.or))
round(OR.CI, 4)

rev(round(1/OR.CI, 4))
```

The odds ratio says that the odds of a successful kick is 0.303 times as large after a time-out than if there was no time out. The confidence interval states with 95% certainty that a successful kick after a time-out is between 0.0697 and 1.3173. This is a very large interval so it is hard to tell if it is accurate. 

# Page 54, Number 17, part e

Based on the tests done, the relative risk, and the odds ratio. I would say that the chances of a field goal after a timeout is not different than if no time out was called. I believe this because the chi square test believes there is no difference between the two. While the relative risk and odds ratio said there could be some evidence that a time-out is effective in icing a kicker, the confidence intervals were so wide that we cannot say with certainty that there is an effect. So, based on the evidence, I don't believe there is a relationship.

# Page 54, Number 18, part a
```{r}
d.table <- array(data = c (135, 15, 434, 9) , dim = c(2, 2),
dimnames = list("Condom" = c("Never", "Ever"),
"HIV" = c("Positive", "Negative")))
d.table
```

```{r}
#Wald
pi.hat.table <- c.table/rowSums(d.table)
alpha <- 0.05
pi.hat1 <- pi.hat.table[1,1]
pi.hat2 <- pi.hat.table[2,1]
var.wald <- pi.hat1 * (1 - pi.hat1)/sum(d.table[1 ,]) +
pi.hat2 * (1 - pi.hat2) / sum(d.table [2 ,])
pi.hat1 - pi.hat2 + qnorm(p = c(alpha / 2, 1- alpha / 2))*
sqrt(var.wald)
```

In the Wald confidence interval, the interval does not contain a 0, which means there is sufficient evidence to indicate a difference in using a condom.

```{r}
#Agresti-Caffo
pi.tilde1 <- (d.table[1 ,1] + 1)/(sum(d.table[1 ,]) + 2)
pi.tilde2 <- (d.table[2 ,1] + 1)/(sum(d.table[2 ,]) + 2)
var.AC <- pi.tilde1 * (1 - pi.tilde1)/(sum(d.table [1,]) + 2) + pi.tilde2 * (1 - pi.tilde2)/(sum(d.table [2 ,]) + 2)
pi.tilde1 - pi.tilde2 + qnorm(p = c(alpha / 2, 1-alpha /2))*sqrt(var.AC)
```

For the Agresti-Caffo confidence interval, the interval also does not contain a 0, which means there is sufficient evidence to indicate a difference in using a condom either. Both models agree, so that means it really does matter if you use a condom, statistically.

# Page 54, Number 18, part b
```{r}
library(package = PropCIs)
prop.test(x = d.table, conf.level = 0.95, correct = FALSE)
```

```{r, warning=FALSE}
chisq.test(x = d.table, correct = FALSE)
```

# Page 54, Number 18, part c
```{r}
pi.hat.table <- d.table/rowSums(d.table)
pi.hat.table
```

```{r}
pi.hat1 <- pi.hat.table[1,1]
pi.hat2 <- pi.hat.table[2,1]

round(pi.hat1/pi.hat2, 4)

round(1/(pi.hat1/pi.hat2), 4) #inverted

alpha<-0.05
n1 <- sum(d.table[1,])
n2 <- sum(d.table[2,])
var.log.rr <- (1-pi.hat1) / (n1*pi.hat1) + (1-pi.hat2) / (n2*pi.hat2)
var.log.rr
```

```{r}
ci <- exp(log(pi.hat1/pi.hat2) + qnorm(p = c(alpha/2, 1-alpha/2)) * sqrt(var.log.rr))
round(ci, 4)

rev(round(1/ci, 4)) #inverted
```

The relative risk inverted is 2.6343, which means it is 260% more likely that a condom will help prevent HIV. With 95% confidence, we can say a condom helps prevent HIV 180% to 371% more effectively than if one isn't used. This interval shows that using a condom is extremely better at preventing HIV than if one isn't used.
